---
title: "scIB Results"
output:
  rmarkdown::html_document:
    toc: false
    number_sections: false
params:
  datasets:
    - Dataset1
    - Dataset2
  methods:
    - Method1
    - Method2
---

```{r setup, include=FALSE}
source(here::here("R", "document_setup.R"))
```

This website displays additional results from the single-cell integration benchmarking project. 
Here you can view all the individual benchmarking metrics broken down by dataset and method.

<div class = "row">
<div class = "col-md-2"></div>
<div class = "col-md-3">

## Datasets

```{r datasets}
dataset_list <- paste(
    glue::glue(
        "* [{params$datasets}](dataset_{params$datasets}.html)"
    ),
    collapse = "\n"
)
```

`r dataset_list`

</div>
<div class = "col-md-2"></div>

<div class = "col-md-3">

## Methods

```{r methods}
methods_list <- paste(
    glue::glue(
        "* [{params$methods}](method_{params$methods}.html)"
    ),
    collapse = "\n"
)
```

`r methods_list`

</div>
<div class = "col-md-2"></div>
</div>

For more information please refer to the main [scIB GitHub page][scIB-github] or the [preprint][preprint].
You can also view the code for this website on [GitHub][scIB-results-github].

If any part of the project is useful for your work please cite:

> Luecken MD, Buttner M, Chaichoompu K, Danese A, Interlandi M, Mueller MF, Strobl DC, Zappia L, Dugas M, Colome-Tatche M, Theis FJ.
> _"Benchmarking atlas-level data integration in single-cell genomics"_
> **bioRxiv**. 2020
> DOI: [10.1101/2020.05.22.111161][preprint]

[scIB-github]: https://github.com/theislab/scib "scIB GitHub repository"
[scIB-results-github]: https://github.com/theislab/scib "scIB results GitHub repository"
[preprint]: https://doi.org/10.1101/2020.05.22.111161 "scIB preprint"
