---
title: "Usability"
---

```{r setup, include=FALSE}
source(here::here("R", "document_setup.R"))
knitr::opts_chunk$set(fig.path = here("..", "docs", "figures", "usability"))
```

```{r libraries}
library("reactable")
library("htmltools")
```

```{r functions}
source(here("R", "tables.R"))
```

This page displays usability scores for each of the methods based on the
information presented in associated papers, package documentation and GitHub
activity.

```{r load}
usability <- drake::readd("usability")
```

:::: {.noticebox}
<!-- :::{.center} -->
**Note about usability scores**
<!-- ::: -->

The usability data was collected prior to 17 December 2020 and may have changed since.
Please refer to the various packages, documentation and publications for the most up-to-date information.
::::

Paper scores
============

These scores assess how well publications associated with a method explain how it works and demonstrate its effectiveness.

* **Peer-reviewed** - The method is published in a peer-reviewed journal
* **Evaluation of accuracy**
  * **Datasets** - The paper shows the method's usefulness on more than two (1.0), two (0.5) or one (0.25) real datasets
  * **Simulations** - The paper quantifies the accuracy of the method using a simulated dataset
* **Robustness** - The paper assessed method robustness (to eg. noise, feature selection, down-sampling, parameter changes, stability) in one (0.25) , two (0.5), or more than two (1.0) ways
* **Benchmarking** - The paper compares the performance of the method against other published methods

```{r papers}
make_papers_table(usability$papers)
```

Package scores
==============

These scores assess the software packages that implement a method.

* **Open source**
  * **Code** - The packages code is freely available
  * **Platform** - The code can be run using a freely available platform or programming language
* **Version control** - The code is available on a public version controlled repository, such as GitHub
* **Unit tests** - The package includes unit tests
* **Tutorial**
  * **Has tutorial** - The package has a vignette or tutorial
  * **Errors** - The steps in the tutorial can be followed without any issues or errors
  * **Scenarios** - The tutorial shows the method's usefulness on more than two (1.0), two (0.5), one (0.25) or no different scenarios
  * **Non-native** - The tutorial shows how to run the method in a non-native language (i.e. R to Python or vice versa)
* **Function documentation**
  * **Purpose** - The purpose and usage of functions/commands is documented
  * **Parameters** - The parameters of functions/commands are documented
  * **Output** - The output of functions/commands is documented
* **GitHub issues**
  * **Activity** - The amount of issues the GitHub repository receives
  * **Response** - How quickly new GitHub issues are responsed to

```{r packages}
make_packages_table(usability$packages)
```
